A Quasi-Genuine FIFO Total Order Multicast Primitive

### 2011-07-14 textos/optpaxos/draftv3.tex

\begin{figure} 
  \centering
  \includegraphics[width=0.8\linewidth]{images/optdel}
  \caption{Execution examples when the optimistic assumption holds (a) and when it fails to hold (b)}
  \label{fig:optdel}
\end{figure}

@@

@@

This property means that, if a message $m$ was delivered to process $p$ of group $G$, then it has (1) been multicast before, (2) $G \in m.dst$ and (3) it has not been delivered by $p$ before.

For a message $m$ to be delivered (l. \ref{algline:consdeliver}), it must be contained in the $barPending$ list (from the condition in line \ref{algline:inbarpending}).  There are two possibilities to when $m$ has been inserted into such list:
\begin{itemize}
  \item $m$ has been originated in the same same group of $p$, that is, $p \in m.dst$, which means $m$ was inserted into $barPending$ in line \ref{algline:insbp2} of Algorithm \ref{algorithm:deliveryminimal}, or
  \item $m$ has been sent from a group $G \neq G'$, which means that it was inserted into $barPending$ in line \ref{algline:insbp1}. 
\end{itemize}

@@

Finally, as for Algorithm \ref{algorithm:nullperiodic}, $null$ messages are never delivered, so they will never violate the uniform integrity property.

\begin{flushright}
$\square$
\end{flushright}

>>>>

\begin{figure} 
  \centering
  \includegraphics[width=0.8\linewidth]{images/optdel}
  \caption{Execution examples when the optimistic assumption holds (a) and when its failure to hold causes a violation of the timestamp order (b)}
  \label{fig:optdel}
\end{figure}

@@

%===           ====================================
%===   LEMMA   ====================================
%===           ====================================
\begin{lems} \label{lemma:mcastbarpending}
Once a message has been \cms{}, it will be inserted into the $barPending$ list of all the correct processes to which such message has been addressed.
\end{lems}

\begin{proof}
In Algorithm \ref{algorithm:deliveryminimal}, whenever a process in a group $G$ \cms{}s a message, it is first \rmc{} to all processes in $G$ (l. \ref{algline:rmlocal}). As the \rmc{} is reliable, every correct process from $G$ \rmd{}s $m$. Once $m$ is \rmd{}ed by each process $p \in G$, it is inserted into the $propPending$ list of that process, to be included in the next proposal from $p$. As every process from $G$ received $m$ and enqueued it for being proposed, and since $m$ is only removed from $propPending$ when decided (l. \ref{algline:removedecided}), all processes will proposed it at some point and $m$ will eventually be decided. At this point, if $G \in m.dst$, every correct process of $G$ will insert $m$ into its $barPending$ list (l. \ref{algline:insbp2}). Besides, each correct process in $G$ \rmc{}s $m$ to all other destination groups that $m$ might be adressed to.

Let $p'$ be any process in a destination group $G'$ of the message $m$, such that $G' \neq G$. In this case, once $m$ is \rmd{}ed at $p'$, it inserts such message into the $barPending$ list, unless it has already done so (lines \ref{algline:notbptwice} and \ref{algline:insbp1}).
\end{proof}

%===           ====================================
%===   LEMMA   ====================================
%===           ====================================
\begin{lems} \label{lemma:barrierperiodic}
  Given a message $m$ in the $barPending$ list of a correct process $p$, from group $G$, eventually every group $G' \in sendersTo(G)$ will have sent $barrier(G',G) > m.ts$ to $p$.
\end{lems}

\begin{proof}Every process $q$ keeps executing Algorithm \ref{algorithm:nullperiodic}, thus sending ever increasing barrier values to the processes in $receiversFrom(group(q))$. From the given definitions, it is obvious that if $G \in receiversFrom(G')$, then $G' \in sendersTo(G)$. Besides, once decided and inserted into the $barPending$ list, the timestamp of the messages no longer can change. This means that, at some point, all of the barrier values received by $p$ will be greater than $m.ts$.
\end{proof}

%===           ====================================
%===   LEMMA   ====================================
%===           ====================================
\begin{lems} \label{lemma:groupfifo}
If a process $p$ from a group $G$ has received a message $m$ from another group $G' \neq G$, then $p$ has also received any message $m'$ from $G'$, where $m'.ts < m.ts$.
\end{lems}

\begin{proof}
Every process $p'$ from $G'$ sends messages to the processes of $G$, including $p$, using a FIFO channel, via the \rmc{} primitive, in the same order in which messages are decided within $G'$. Also, the agreement property of consensus ensures that all processes from $G'$ will decide the same messages in each consensus instance $k$. Finally, $p'$ \rmc{}s to $G$ every message $m : m.src = G' \wedge G \in m.dst$ (l. \ref{algline:rmothers} of Algorithm \ref{algorithm:deliveryminimal}).

Assume, by way of contradiction, that $p$ received a message $m':m'.ts < m.ts$ from some process $p'' \in G'$ after having received $m$ from $p'$. As $m'.src = m.src$, this means that $p$ either will send $m'$ afterwards, contradicting the fact that the processes in $G'$ agree on each batch of messages decided and send them to $G$ in the same FIFO order, or that $p'$ will not send $m'$ at all, contradicting the fact the the \rmc{} primitive is reliable. Therefore, the assumption is false and the lemma is valid.
\end{proof}

%===           ====================================
%===   LEMMA   ====================================
%===           ====================================
\begin{lems} \label{lemma:alwaysbpdel}
Once a message has been inserted into either the $barPending$ list or the $delivered$ list of a given process, it will always belong to one of them.
\end{lems}

\begin{proof}
Immediate from the algorithm, considering that a message is never removed from the $delivered$ list, and that it is removed from $barPending$ (l. \ref{algline:rmfrombp}) right before being inserted into $delivered$ (l. \ref{algline:addtodelivered}).
\end{proof}

%===           ====================================
%===   LEMMA   ====================================
%===           ====================================
\begin{lems} \label{lemma:oncedecidednotsmaller}
  Once a process inserts a message $m$ from some group into its $barPending$ list, no message $m' : m'.ts < m.ts$ from the same group will be inserted into the same $barPending$ list afterwards.
\end{lems}

\begin{proof}
Let $p$ be a process from the group $G$ which sent $m$. Before inserting $m$ into the $decided$ list (l. \ref{algline:addtodecided} of Algorithm \ref{algorithm:deliveryminimal}), $p$ checks whether there is already some other message with an equal or lower timestamp in such list. If that is the case, the timestamp of $m$ is changed to a value greater than that of any other message already within $decided$. Only then, if $G$ is a destination of $m$, $p$ inserts $m$ into its $barPending$ list (l. \ref{algline:insbp2}). Therefore, for processes from the group from which $m$ was sent, it is inserted in the $barPending$ list always in ascending order of timestamp. Besides, each process from $G$ \rmc{}s $m$ to its other destination groups in this same order (l. \ref{algline:rmothers}).

Regarding groups other than the source of the message, let $p'$ be any process that is a destination of $m$, but does not belong to the group $G$ from where $m$ came. The process $p'$ inserts $m$ into its $barPending$ list in line \ref{algline:insbp1}. Considering that every process in a group sends that group's messages to other groups in the same order in which they are decided, that once a group has received some message from another group, any message prior to that one from the same group has also already been received (from Lemma \ref{lemma:groupfifo}), and that no message is inserted twice in the $barPending$ list of any process (from line \ref{algline:notbptwice} and Lemma \ref{lemma:alwaysbpdel}), we can infer that, once $p'$ inserts $m$ into its $barPending$ list, no message $m' : m'.ts < m.ts$ from $G$ will be inserted into the same $barPending$ list afterwards.
\end{proof}

%===           ====================================
%===   LEMMA   ====================================
%===           ====================================
\begin{lems} \label{lemma:oncebpwilldel}
Once a message $m$ has been inserted into the $barPending$ list of a process $p$, it will eventually be removed from such list and, if $m$ is different from $null$, it will be delivered by $p$.
\end{lems}

\begin{proof}
Eventually, $p$ will have received a barrier from every group in $sendersTo(group(p))$ with a value greater than the timestamp of $m$ (from Lemma \ref{lemma:barrierperiodic}). Once this happens, as any barrier is also a message, no more messages with a timestamp lower than that of $m$ will be inserted into the $barPending$ list of $p$ (from Lemma \ref{lemma:oncedecidednotsmaller}). Let $barReady$ be the set of $m$ plus all messages from $barPending$ whose timestamps are lower than that of $m$, that is, $barReady = \{m\} \cup \{m' : m' \in barPending \wedge m'.ts < m.ts\}$.

It is important to note that, from Lemma \ref{lemma:oncedecidednotsmaller}, no more messages will be included in this set, meaning that any other message that might be later inserted into $barPending$ will have a timestamp greater than that of any message in $barReady$. Besides, the barrier received for $m$ also apply to all the other messages in this set, since their timestamps are lower than that of $m$. We can, therefore, prove that all these messages will eventually be delivered by induction on the position $i$ of each message $m_i$ in $barReady$, in ascending order of timestamps.\\
\\
\noindent Base case ($i=1$): Let $m_1$ be the first message in $barReady$. It means that $\nexists m \in barPending : m.ts < m_1.ts$. The message $m_1$ has the lowest timestamp in the list and all the necessary barriers have been received already. Therefore, $m_1$ satisfies all conditions from line \ref{algline:inbarpending} of Algorithm \ref{algorithm:deliveryminimal} and will be removed from $barPending$. If it is different from $null$, it will also be delivered.\\
\\
\noindent Induction step: Suppose that $m_{i}$ will eventually be removed from $barPending$. Once this happens, it means that $m_{i}$ was the first message in the list. Since no more messages have been inserted into $barReady$, as soon as $m_{i}$ is removed, $m_{i+1}$ will be the new first one. Then, as $m_{i+1}$ will have the lowest timestamp in $barReady$ -- and, therefore, in $barPending$ --, it will satisfy both conditions of line \ref{algline:inbarpending} from Algorithm \ref{algorithm:deliveryminimal}), being removed from $barPending$ -- and delivered, if it is different from $null$. \end{proof}

%===                 ==============================
%===   PROPOSITION   ==============================
%===                 ==============================
\begin{props}
Once a process \cms{}s a message, then one of the correct processes that is a destination of that message eventually delivers it.
\end{props}

\begin{proof}
Immediate from Lemma \ref{lemma:mcastbarpending}, Lemma \ref{lemma:oncebpwilldel}, and from the assumption that the application will not \cms{} a $null$ message -- not in the \cms{} protocol level at least; if the application needs to \cms{} an empty message, in the protocol level it would not be handled as a $null$ message.
\end{proof}

@@

%===                 ==============================
%===   PROPOSITION   ==============================
%===                 ==============================
\begin{props} \label{props:integrity}
If a message $m$ was delivered to process $p$ of group $G$, then it has (1) been \cms{} before, (2) $G \in m.dst$ and (3) it has not been \cons{}ed by $p$ before.
\end{props}

\begin{proof}
For a message $m$ to be delivered (l. \ref{algline:consdeliver}), it must be contained in the $barPending$ list (from the condition in line \ref{algline:inbarpending}).  There are two possibilities to when $m$ has been inserted into such list:
\begin{itemize}
  \item $m$ has been originated in the same same group of $p$, that is, $p \in m.dst$, which means $m$ was inserted into $barPending$ in line \ref{algline:insbp2} of Algorithm \ref{algorithm:deliveryminimal}, or
  \item $m$ has been sent from a group $G \neq G'$, which means that it was inserted into $barPending$ in line \ref{algline:insbp1}. 
\end{itemize}

@@

Finally, as for Algorithm \ref{algorithm:nullperiodic}, $null$ messages are never delivered, so they will never violate the uniform integrity property.
\end{proof}

%===                 ==============================
%===   PROPOSITION   ==============================
%===                 ==============================
\begin{props}
If a correct process which is a destination of a message $m$ delivers it, then all correct processes that are also destinations of $m$ deliver it as well.
\end{props}

\begin{proof}
Immediate from Proposition \ref{props:integrity}, Lemma \ref{lemma:mcastbarpending} and Lemma \ref{lemma:oncebpwilldel}.
\end{proof}
